\section{Introduction}

The FDA came about as the first consumer protection agency in the US. Since its inception, it has carried out the mission of protecting the public health by ensuring the safety, efficacy, and security of human and veterinary drugs, biological products, and medical devices; and by ensuring the safety of our nation's food supply, cosmetics, and products that emit radiation. The FDA also has a role in advancing public health by helping to speed innovations that make medicines and foods more effective, safer, and more affordable; and by helping the public get the accurate, science-based information they need to use medicines and foods to improve their health.

However, the mission has not been without controversy. The FDA has been criticized for being too slow to approve new drugs and medical devices, which can delay access to potentially life-saving treatments. On the other hand, it has also been criticized for being too fast in approving drugs that later turn out to be unsafe or ineffective. The FDA's role in regulating food safety has also been a source of controversy, with some arguing that it is not doing enough to protect consumers from foodborne illnesses.

Even still, the FDA remains a crucial agency in ensuring the safety and efficacy of products, namely of drugs, cultivating over its 100-year history a reputation for competence and vigilance against unsafe drugs---more than its de facto legal authority---even as administrations have come and gone \cite{carpenterINTRODUCTIONGatekeeper2010}.

Beyond its domestic role, the FDA has also become one of the most influential pharmaceutical regulators in the world. Because access to the U.S. pharmaceutical market is economically critical for drug manufacturers, regulatory decisions made by the FDA often shape the global development and approval strategies of pharmaceutical firms. As a result, the agency is frequently regarded as a benchmark for regulatory rigor, and regulatory agencies in other countries often observe, emulate, or coordinate with FDA standards and practices. [CITE GLOBAL FDA INFLUENCE]

Despite this reputation for scientific rigor and regulatory vigilance, the FDA's drug approval process has long been subject to policy debate. Critics have argued that lengthy approval times delay access to beneficial treatments, while others warn that accelerating approvals may increase the risk that unsafe or ineffective drugs reach the market. These tensions highlight a central policy tradeoff in pharmaceutical regulation: balancing the benefits of faster access to new therapies against the potential risks to patient safety. [CITE ECON LITERATURE ON FDA SPEED VS SAFETY]

One of the most significant policy changes addressing this tradeoff was the Prescription Drug User Fee Act (PDUFA) of 1992. PDUFA allowed the FDA to collect user fees from pharmaceutical manufacturers to fund the drug review process, with the goal of accelerating regulatory review times and bringing new drugs to market more quickly. A large literature has examined the consequences of this policy, focusing primarily on how faster approval times affect drug safety, innovation incentives, and welfare outcomes. [CITE PDUFA WELFARE LITERATURE]

However, relatively little attention has been paid to how changes in approval speed may affect the
*composition* of drugs that reach the market. If expedited review mechanisms disproportionately benefit certain types of pharmaceutical products—particularly those with high expected market returns—then regulatory reforms designed to accelerate approval could alter not only the number of drugs approved, but also the types of drugs entering the market. In particular, pharmaceutical firms producing highly profitable drug classes may have stronger incentives to utilize expedited regulatory pathways.

This paper investigates whether policies designed to accelerate FDA drug approvals have changed the composition of approved drugs in the United States. Specifically, it examines whether the introduction of PDUFA is associated with an increase in the approval of controlled substances relative to other pharmaceutical products. Because controlled substances—such as opioid analgesics—often generate substantial market revenues while also posing significant public health risks, understanding how regulatory institutions shape their entry into the market is of particular policy relevance.

More broadly, examining how regulatory reforms affect the composition of drug approvals can shed light on the economic incentives embedded within pharmaceutical regulation. If policies intended to improve regulatory efficiency systematically favor certain categories of drugs, then regulatory institutions may influence not only the speed of innovation, but also the direction of pharmaceutical development.
